### 2019-10-21 Common/chapters/technical_info.tex
Julien and Johan: more details about DICOM and paragraph about RAW and MHD

% !TEX root = ../main.tex

\chapter{Technical information}
\label{ch:technical_info}

\section{DICOM}

\subsection{Origin}

The acronym DICOM stands for Digital Imaging and Communications in Medicine. Before the 1980’s, images resulting from CT scans and MRIs were only decodable by machine manufacturers, while the medical community needed to export and share them for other tasks. For that reason, the ACR (American College of Radiology) and the NEMA (National Electrical Manufacturers Association) created a committee to build a standard. After two iterations with other names, DICOM was created in 1993. It standardized the representation of medical images and their transmission since it provided a network protocol built on top of TCP/IP.

\subsection{Data format}

DICOM files can be viewed as containers of attributes, also called tags. The values of the pixels themselves are stored under the "Pixel Data" tag. Every single DICOM file usually represents a 2-dimensional image, which will form a 3-dimensional volume when put all together.

Other useful information such as the patient name and ID is directly stored within the DICOM files. This approach aims at linking each image to a specific person and event in order not to mix them up. Each DICOM file can be seen as part of a bigger dataset.

\subsection{Processing images}

When manipulating DICOM files, multiple details must be taken into account.

\subsubsection{Order}
First of all, the name of the files within datasets is a 6 digit number, from 000000 to the number of images minus one. However, this order doesn’t match the real order of the images. In fact, the correct order is given by the "Instance Number" tags contained in the various files. Therefore, converted images must be sorted by instance number.

\subsubsection{Data manipulation}
CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other. DICOM takes this problematic into account by providing specific tags that allow to display the exact same representation of the data, no matter the hardware used. Otherwise, physicians may struggle to detect anomalies because of color and exposition-related variations. 
Therefore, before displaying or converting an image to any format (such as png), pixel data must be normalized.

The procedure depends on the tags "Window Width" and "Window Center" (one always come with the other). These are used to represent a range of values corresponding to the pixel values in the data. For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.

If they are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
\begin{itemize}
	\item Bits allocated: the number of bits used to represent a single pixel (value: 1 or a multiple of 8)
	\item Samples per pixel: the number of channels for each pixel

\end{itemize}

\noindent Examples: 
\begin{itemize}
\item 1 bit, 1 sample: black and white
\item 8 bits, 1 sample: grayscale
\item 8 bits, 3 samples: RGB
\item 16 bits, 1 sample: grayscale

\end{itemize}

If they are included in the DICOM header, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one. To do this, two step are required:

\begin{enumerate}
	\item \textbf{Apply the Hounsfield correction}\newline
	Hounsfield Units (HU) are used in CT images it is a measure of radio-density, calibrated to distilled water and free air. Provided that the rescale slope and the rescale intercept are included in the DICOM header, the correction is applied thanks to the following formula:
\begin{equation}
	HU = m * P + b
\end{equation}
	where~$m$ is the rescale slope,~$P$ the pixel value, ~$b$ the rescale intercept.

	\item \textbf{Apply a linear transformation}\newline
	The result of the first operation then goes through a linear transformation based on the following conditions:

\begin{equation}\label{eq:dicom1}
\textrm{if } (P \leq c - 0.5 - \frac{w-1}{2}) \textrm{, then }y = y_{min}
\end{equation}

\begin{equation}\label{eq:dicom2}
\textrm{else if } (P > c - 0.5 + \frac{w-1}{2}) \textrm{, then }y = y_{max}
\end{equation}

\begin{equation}\label{eq:dicom3}
\textrm{else } y = (\frac{P - (c - 0.5)}{w-1}) + 0.5) * (y_{max} - y_{min}) + y_{min}
\end{equation}

where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, ~$y$ the pixel output value, ~$y_{min}$ the minimal value of the output range (usually 0), ~$y_{max}$ the maximal value of the output range (usually 255). Equations \ref{eq:dicom1}, \ref{eq:dicom2} and \ref{eq:dicom3} ensure that the pixel values are correctly distributed within the output range.

\end{enumerate}

\section{NIfTI}

\subsubsection{Origin}
The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format. The main problem of the latter was that it was lacking information about orientation in space. Therefore, the interpretation of stored data could be problematic and inconsistent. For instance, there was a real confusion to determine the left and right sides of brain images. Hence, the NIfTI file format was defined to overcome this major issue.

\subsubsection{Data format}
Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility. This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data. The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate”  algorithm.

\subsubsection{Overview of the header structure}
With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes. “Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”. Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}

\section{RAW and MHD}

Some datasets use a combination of RAW and MHD files. The latter contain metainformation about their corresponding RAW file(s) which contain the data. In most cases, each MHD file points to a unique RAW file whose name is the same as the MHD file name. A single RAW file can be used to represent three-dimensional data, i.e. the combination of multiple two-dimensional images. Libraries such as \mbox{SimpleITK} in Python allow to manipulate RAW images in an easy way.

\section{Conversion}

Medical data is often represented over 16 bits. However, exporting images to PNG require 8-bit images. To transpose a 16-bit image to an 8-bit one, the following procedure was applied:

\begin{enumerate}
	\item Find maximum value of the current image $Pixel_{min}$
	\item Find minimum value of the current image $Pixel_{max}$
	\item Pixel normalization: $Pixel = \frac{(Pixel - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
	\item Modify object type to 8 bits
	\item Export to PNG
\end{enumerate}