### 2020-03-04 Common/structure/title.tex

% !TEX root = ../main.tex

%----------------------------------------------------------------------------------------
% TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

%\includegraphics[width=15cm]{logos/xi_logos.pdf}
%\vspace*{.06\textheight}
%\\

\begin{figure}
  \centering
    \includegraphics[width=.17\textwidth]{logos/unifr_logos.pdf}
  \hfill
    \includegraphics[width=.4\textwidth]{logos/xi_logos.pdf}
  \vspace{30mm}
\end{figure}

{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
\textsc{\Large Master's Thesis}\\[0.5cm] % Thesis type
\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Authors:}\\
{\authorname} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
\href{https://exascale.info/phil}{\supname} % Supervisor name - remove the \href bracket to remove the link
\\\vspace*{1ex}\emph{Co-Supervisors:} \\ % Remove these two lines if no co-supervisor is involved
{\cosupname} % Co-supervisor name - remove the \href bracket to remove the link
\end{flushright}
\end{minipage}\\[1cm]

March XX, 2020 % date of the official defense
\vspace*{.06\textheight}

%\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\textit{ in the}\\[0.3cm] % University requirement text
\groupname\\\deptname\\ % Research group name and department name
\vfill

\footnotesize{ Boulevard de Pérolles 90 ~~$\bullet$~~ 1700~Fribourg ~~$\bullet$~~ Switzerland
            \\
            phone +41~(26)~300~84~65 ~~$\bullet$~~ \textsf{diuf-secr@unifr.ch} ~~$\bullet$~~ \textsf{www3.unifr.ch/inf}
            }

\end{center}
\end{titlepage}