### 2020-03-04 Common/chapters/acknowledgements.tex
Johan and Julien: Acknowledgements

\textbf{Common acknowledgements:}\\
First of all, we would like to thank Giuseppe Cuccu, for his valuable advices, the exciting discussions we had with him and all the time he took to supervise the project. Even when the results were not there and we were struggling to find the cause, he always encouraged us and pushed us to give our best. We are extremely grateful to him. We also would like to greatly thank Akansha Bhardwaj for her supervision throughout the project and the amazing work she did reviewing the thesis. Last but not least, we would like to express sincere thanks to Philippe Cudré-Mauroux for having supervised the project. Having the opportunity to work with such a team and having access to amazingly powerful machines was literally awesome.\\

\noindent \textbf{Julien Clément's acknowledgements:}\\
XXX\\

\noindent \textbf{Johan Jobin's acknowledgements:}\\
On a personal note, I would like to acknowledge with gratitude the support of my family, specially my parents (Jean-Bernard and Marie Jobin), my brother (Lucas Jobin), my uncle (Xavier Marchand) and my girlfriend (Julie Timmermans). They have allowed me to be in the best conditions to carry out this work and have expressed great interest in it, which gave me even more motivation. A special thanks goes to my close friends that have always encouraged me. Finally, I would like to thank Julien Clément for the great work we did together on this thesis but also for all the years we have spent at the university. The realization of this work would not have been possible without him.