### 2020-03-06 Common/structure/info.tex

% !TEX root = ../main.tex

%----------------------------------------------------------------------------------------
% THESIS INFORMATION
%----------------------------------------------------------------------------------------

\thesistitle{Prostate Cancer Classification:\\A Freeze/Unfreeze Transfer Learning Approach Using Multiple Body Parts} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Prof. Dr. Philippe Cudré-Mauroux} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\cosupervisor{\href{https://exascale.info/members/giuseppe-cuccu/}{Dr. Giuseppe Cuccu},\\ \href{https://exascale.info/members/akansha/}{Akansha Bhardwaj}} % If you have a co-supervisor, include it here. This is used in the title page, print it elsewhere with \supname
\degree{Master} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{\href{mailto:julienclement22@gmail.com}{Julien Clément}\\ \href{mailto:joh.job13@gmail.com}{Johan Jobin}}% Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{Bd de Pérolles 90} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{Computer Science} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{Prostate cancer classification, Convolutional Neural Network (CNN), Transfer learning, PROSTATEx, Machine Learning (ML), Deep Learning (DL), Artificial Intelligence (AI)} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{http://www.unifr.ch}{University of Fribourg}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\department{\href{https://www3.unifr.ch/inf/fr/}{Department of Informatics}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\group{\href{https://www3.unifr.ch/inf/en/exascale-infolab.html}{eXascale Infolab}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\faculty{\href{https://www3.unifr.ch/scimed/fr/}{Faculty of Science and Medicine}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname